Do we need detailed demographic data to forecast plant population responses to climate change?

### 2014-12-17 manuscript/manuscript.Rmd
ms ref updates

bibliography: components/references.bib
csl: components/ecology.csl
documentclass: components/elsarticle

@@

Introduction
============
Producing skillfull forecasts at spatial scales relevant to global environmental changes that play out at the landscape and regional scale has proven challenging for population ecologists. The major hurdle is that most population models, at least for plant species, are built using data from small, localized plots because parameterizing vital rate regressions requires tracking the fates of individuals. These models are difficult to scale up from the micro to meso-scales because the fitted parameters do not fully represent the spatial variation present at spatial scales beyond that at which the data are collected.

Recently, @Freckleton2011, building on work by Hastings and NAME, have proposed density-structured population models that focus on the transition of populations among discrete states, rather than the traditional approach of modeling the transitions of individuals. Such an approach could be extremely valuable because the data needed to parameterize density-structured population models is much easier, and less costly, to collect (Queenborough et al. year). For example, using a density-structured approach, one could build a population model using a time series of annual plot-based censuses of species percent cover. However, a major assumption of the density-structured approach is that the aggregate dynamics of the population observed at coarse spatial resolution faithfully represent, and correspond to, the fates of individual plants. In other words, using a density-structured approach requires a leap of faith that important covariates (e.g., climate variables) at the level of the individual are captured adequately at the population level. If we seek to forecast the impacts of climate change on plant populations, then clearly this assumption requires testing.

@@

Results
=======
We assessed the statistical importance of the climate covariates included the final vital rate regressions by comparing the residual deviance of models with climate covariates and temporal random effects, climate covariates only, and temporal random effects only. When a model includes climate covariates, this comparison shows the relative contribution of the climate covariates in explaining the total interannual variability (Adler et al. 2012).

@@

>>>>

bibliography: ~/Dropbox/Bibliography/MicroMeso.bib
csl: components/ecology.csl
documentclass: components/elsarticle

@@

Introduction
============
Population models are important tools for predicting the impacts of environmental change on species. But reconciling the scales at which population models are parameterized and the scales at which environmental changes play out remains a challenge [@Freckleton2011; @Queenborough2011]. The major hurdle is that most population models, at least for plant species, are built using data from small, localized plots because parameterizing traditional population models requires tracking the fates of individuals. These models are difficult to scale up from the micro to meso-scales because the fitted parameters do not fully represent the spatial variation present at scales beyond that at which the data are collected [@Sæther2007]. Thus, our ability to use population models to predict the consequences of climate change is limited when we rely on individual-level data.

In contrast to individual-level data, population-level data arising from large-scale census efforts

Recently, @Freckleton2011, building on work by @Taylor2004, have proposed density-structured population models that focus on the transition of populations among discrete states, rather than the traditional approach of modeling the transitions of individuals. Such an approach could be extremely valuable because the data needed to parameterize density-structured population models is much easier, and less costly, to collect.

For example, using a density-structured approach, one could build a population model using a time series of annual plot-based censuses of species percent cover. However, a major assumption of the density-structured approach is that the aggregate dynamics of the population observed at coarse spatial resolution faithfully represent, and correspond to, the fates of individual plants. In other words, using a density-structured approach requires a leap of faith that important covariates (e.g., climate variables) at the level of the individual are captured adequately at the population level. If we seek to forecast the impacts of climate change on plant populations, then clearly this assumption requires testing.

@@

Results
=======
We assessed the statistical importance of the climate covariates included the final vital rate regressions by comparing the residual deviance of models with climate covariates and temporal random effects, climate covariates only, and temporal random effects only. When a model includes climate covariates, this comparison shows the relative contribution of the climate covariates in explaining the total interannual variability (@Adler2012).

@@

----------

### 2014-12-17 manuscript/manuscript.tex
new readme for subdirectories
ms ref updates

Producing skillfull forecasts at spatial scales relevant to global
environmental changes that play out at the landscape and regional scale
has proven challenging for population ecologists. The major hurdle is
that most population models, at least for plant species, are built using
data from small, localized plots because parameterizing vital rate
regressions requires tracking the fates of individuals. These models are
difficult to scale up from the micro to meso-scales because the fitted
parameters do not fully represent the spatial variation present at
spatial scales beyond that at which the data are collected.

Recently, Freckelton et al., building on work by Hastings and NAME, have
proposed density-structured population models that focus on the
transition of populations among discrete states, rather than the
traditional approach of modeling the transitions of individuals. Such an
approach could be extremely valuable because the data needed to
parameterize density-structured population models is much easier, and
less costly, to collect (Queenborough et al. year). For example, using a
density-structured approach, one could build a population model using a
time series of annual plot-based censuses of species percent cover.
However, a major assumption of the density-structured approach is that
the aggregate dynamics of the population observed at coarse spatial
resolution faithfully represent, and correspond to, the fates of
individual plants. In other words, using a density-structured approach
requires a leap of faith that important covariates (e.g., climate
variables) at the level of the individual are captured adequately at the
population level. If we seek to forecast the impacts of climate change
on plant populations, then clearly this assumption requires testing.

@@

We assessed the statistical importance of the climate covariates
included the final vital rate regressions by comparing the residual
deviance of models with climate covariates and temporal random effects,
climate covariates only, and temporal random effects only. When a model
includes climate covariates, this comparison shows the relative
contribution of the climate covariates in explaining the total
interannual variability (Adler et al. 2012).

@@

>>>>

Population models are important tools for predicting the impacts of
environmental change on species. But reconciling the scales at which
population models are parameterized and the scales at which
environmental changes play out remains a challenge (Freckleton et al.
2011, Queenborough et al. 2011). The major hurdle is that most
population models, at least for plant species, are built using data from
small, localized plots because parameterizing traditional population
models requires tracking the fates of individuals. These models are
difficult to scale up from the micro to meso-scales because the fitted
parameters do not fully represent the spatial variation present scales
beyond that at which the data are collected (Sæther et al. (2007)).
Thus, our ability to use population models to predict the consequences
of climate change is limited.

Recently, Freckleton et al. (2011), building on work by Hastings and
NAME, have proposed density-structured population models that focus on
the transition of populations among discrete states, rather than the
traditional approach of modeling the transitions of individuals. Such an
approach could be extremely valuable because the data needed to
parameterize density-structured population models is much easier, and
less costly, to collect (Queenborough et al. year). For example, using a
density-structured approach, one could build a population model using a
time series of annual plot-based censuses of species percent cover.
However, a major assumption of the density-structured approach is that
the aggregate dynamics of the population observed at coarse spatial
resolution faithfully represent, and correspond to, the fates of
individual plants. In other words, using a density-structured approach
requires a leap of faith that important covariates (e.g., climate
variables) at the level of the individual are captured adequately at the
population level. If we seek to forecast the impacts of climate change
on plant populations, then clearly this assumption requires testing.

@@

We assessed the statistical importance of the climate covariates
included the final vital rate regressions by comparing the residual
deviance of models with climate covariates and temporal random effects,
climate covariates only, and temporal random effects only. When a model
includes climate covariates, this comparison shows the relative
contribution of the climate covariates in explaining the total
interannual variability (Adler et al. (2012)).

@@

\begin{center}\rule{3in}{0.4pt}\end{center}

\section{References}\label{references}

Adler, P. B., H. J. Dalgleish, and S. P. Ellner. 2012. Forecasting plant
community impacts of climate variability and change: when do competitive
interactions matter? Journal of Ecology 100:478--487.

Freckleton, R. P., W. J. Sutherland, A. R. Watkinson, and S. A.
Queenborough. 2011. Density-structured models for plant population
dynamics. American Naturalist 177:1--17.

Queenborough, S. A., K. M. Burnet, W. J. Sutherland, A. R. Watkinson,
and R. P. Freckleton. 2011. From meso- to macroscale population
dynamics: A new density-structured approach. Methods in Ecology and
Evolution 2:289--302.

Sæther, B. E., S. Engen, V. Grøtan, W. Fiedler, E. Matthysen, M. E.
Visser, J. Wright, A. P. Møller, F. Adriaensen, H. {Van Balen}, D.
Balmer, M. C. Mainwaring, R. H. McCleery, M. Pampus, and W. Winkel.
2007. The extended Moran effect and large-scale synchronous fluctuations
in the size of great tit and blue tit populations. Journal of Animal
Ecology 76:315--325.
